Comparing OTU quality across clustering strategies

### 2021-07-02 exploratory/2021/algorithm_diagram.Rmd
playing around with layouts for algorithm diagram

```{r opticlust}
# example data
seqs <- LETTERS[1:17]
seq_pairs <- combn(seqs, 2)
# distances within threshold
dat <- data.frame(
  Seq1 = c("D", "F", "G", "H", "I", "I", "J", "J", "N", "O", "P", "P", "P", "Q", "Q"),
  Seq2 = c("B", "E", "C", "A", "B", "D", "A", "H", "M", "L", "K", "L", "O", "E", "F"),
  Distance = c(0.024, 0.028, 0.028, 0.027, 0.016, 0.024, 0.028, 0.024, 0.024, 0.024, 0.016, 0.027, 0.027, 0.024, 0.028)
)

# matrix of pairwise sequences within distance threshold
dist_mat <- matrix(
  nrow = length(seqs), ncol = length(seqs),
  dimnames = list(seqs, seqs), data = 0
)
diag(dist_mat) <- 1
for (i in 1:nrow(dat)) {
  r <- dat[i, ]
  seq1 <- r[["Seq1"]]
  seq2 <- r[["Seq2"]]
  dist_mat[seq1, seq2] <- 1
  dist_mat[seq2, seq1] <- 1
}

@@

```{r ggraph_mod}
# devtools::install_github('kelly-sovacool/ggraph', ref = 'iss-297_ggtext')
#library(ggraph) # something is wrong & this fails, but load_all() works!
devtools::load_all('../ggraph')
library(ggtext)
library(tidygraph)
```
```{r ggraph_exp}
graph <-
  tbl_graph(nodes = data.frame(name = c("A B C", "D E F", "G **H** I", "J K L")),
            edges = data.frame(from = c(2, 2, 2), 
                               to = c(3, 4, 2), 
                               mcc = c(0.97, 0.84, 0.80)))
ggraph(graph, layout = 'linear') +
  geom_edge_arc(aes(label = mcc,
                    start_cap = label_rect(node1.name),
                    end_cap = label_rect(node2.name)), 
                arrow = arrow(length = unit(4, 'mm')),
                edge_colour = 'gray',
                angle_calc = 'along',
                label_dodge = unit(2, 'mm')
                ) +
  geom_edge_loop(aes(label = mcc,
                     span=120,
                     direction=270,
                     strength=0.5
                     ),
                 angle_calc = 'along',
                 label_dodge = unit(-2, 'mm'),
                 edge_colour = 'gray') +
  geom_node_label(aes(label = name), richtext = TRUE) +
  theme(panel.background = element_rect(fill='white'))
```

>>>>

```{r opticlust}
# example data
seqs <- LETTERS[1:17]
seq_pairs <- combn(seqs, 2)
# distances within threshold
dist_dat <- data.frame(
  Seq1 = c("D", "F", "G", "H", "I", "I", "J", "J", "N", "O", "P", "P", "P", "Q", "Q"),
  Seq2 = c("B", "E", "C", "A", "B", "D", "A", "H", "M", "L", "K", "L", "O", "E", "F"),
  Distance = c(0.024, 0.028, 0.028, 0.027, 0.016, 0.024, 0.028, 0.024, 0.024, 0.024, 0.016, 0.027, 0.027, 0.024, 0.028)
)

# matrix of pairwise sequences within distance threshold
dist_mat <- matrix(
  nrow = length(seqs), ncol = length(seqs),
  dimnames = list(seqs, seqs), data = 0
)
diag(dist_mat) <- 1
for (i in 1:nrow(dist_dat)) {
  r <- dist_dat[i, ]
  seq1 <- r[["Seq1"]]
  seq2 <- r[["Seq2"]]
  dist_mat[seq1, seq2] <- 1
  dist_mat[seq2, seq1] <- 1
}

@@

```{r ggraph_mod}
# devtools::install_github('kelly-sovacool/ggraph', ref = 'iss-297_ggtext')
#library(ggraph) # something is wrong & this fails, but load_all() works!
devtools::load_all('../ggraph')
library(ggtext)
library(tidygraph)
library(patchwork)
library(gridExtra)
g1 <-
  tbl_graph(nodes = data.frame(name = c("A B C", "D E F", "G **H** I", "J K L"),
                               i = 1:4),
            edges = data.frame(from = c(2, 2, 2), 
                               to = c(3, 4, 2), 
                               mcc = c(0.97, 0.84, 0.80)) %>% 
              mutate(is_loop = from == to))
g2 <-
  tbl_graph(nodes = data.frame(name = c("A B C", "D **E** F", "G H I", "J K L"),
                               i = 1:4),
            edges = data.frame(from = c(2, 2, 2), 
                               to = c(1, 3, 2), 
                               mcc = c(0.94, 0.81, 0.86)) %>% 
              mutate(is_loop = from == to))
```

```{r ggraph_exp, fig.dim=c(6,5), dpi=300}
plot_graph <- function(graph) {
  create_layout(graph, 'linear', sort.by = i) %>% 
  ggraph() +
  geom_edge_arc(aes(label = mcc,
                    start_cap = label_rect(node1.name),
                    end_cap = label_rect(node2.name)), 
                arrow = arrow(length = unit(4, 'mm'),
                              type = 'closed'),
                edge_colour = 'gray',
                angle_calc = 'along',
                label_dodge = unit(-2, 'mm')
                ) +
  geom_edge_loop(aes(span = 1, direction = 270, strength = 0.2, color = is_loop)) +
  geom_node_label(aes(label = name)) +
    scale_edge_color_manual(values = c('white')) +
  theme_void() +
  theme(plot.margin=unit(x=c(0,0,10,0), units="pt"),
        legend.position = 'none')
}

pg1 <- plot_graph(g1) 
pg2 <- plot_graph(g2)

wrap_elements(tableGrob(t(dist_dat),
                        theme = ttheme_default(base_size = 9, 
                                               padding = unit(c(4,8), 'pt')))
              ) /
((pg1 / pg2 / pg1 / pg2) + 
  plot_annotation(tag_levels = '1')) +
plot_layout(heights = c(1,4))
